Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2022-03-06 Thesis/01-chap1.Rmd
add performance

## Artificial intelligence: Four Approaches

@@

Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build. Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research. We believed humanly behaviours was enough for demonstrating AI for the problem we concerned. Hence, adopting one of the capacities required by the Turing test, namely machine learning, along with

Thought process, human visual

## Discussion of potential methods

@@

>>>>

## Artificial intelligence: Four Approaches {#se:ai}

@@

Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build. Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research. We believed humanly behaviours was enough for demonstrating AI for the problem we concerned. Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.

Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions. This method would work if such metrics can be obtained manually or approximated algorithmically. However, we considered it was too complex as a starting point.

## Automatic visual inference system

For an automatic visual inference system, there are several possibilities of the type of inputs.

Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.

@@

```{r}
null_true_eval_dat <- data.frame()

model_name <- c("VGG16", "VGG19", "Xception", "ResNet152V2", "InceptionV3")

for (i in 1:5) {
  mid_dat <- read_csv(here::here(glue::glue("data/cv/prediction/mod_{i}_null_both.csv")))
  null_true_eval_dat <- bind_rows(null_true_eval_dat, mid_dat)
  mid_dat <- read_csv(here::here(glue::glue("data/cv/prediction/mod_{i}_true_both.csv")))
  null_true_eval_dat <- bind_rows(null_true_eval_dat, mid_dat)
} 
```

```{r}
null_true_eval_dat %>%
  group_by(model, null) %>%
  summarise(rate = mean(prediction == null)) %>%
  pivot_wider(names_from = "null", values_from = "rate") %>%
  ungroup() %>%
  mutate(model = model_name) %>%
  mutate(avg_rate = (`FALSE` + `TRUE`)/2) %>%
  knitr::kable(col.names = c("Model", "Actual data plot correct rate", "Null plot correct rate", "Average correct rate"), 
             booktabs = TRUE, 
             caption  = 'Performance of computer vision models on null residual plot and acutal data plot.', 
             label = "null-true-eval" ) 
```

```{r}
lineup_eval_dat <- data.frame()

for (i in 1:5) {
  mid_dat <- read_csv(here::here(glue::glue("data/cv/prediction/mod_{i}_fixed_both.csv"))) %>%
    mutate(model = i) 
  lineup_eval_dat <- bind_rows(lineup_eval_dat, mid_dat)
}

eval_data_para <- readRDS(here::here("data/cv/eval_data_para.rds"))
eval_data_para <- reduce(map(eval_data_para, as.data.frame), bind_rows)
eval_original_data <- readRDS(here::here("data/cv/eval_original_data.rds"))
```

```{r}
eval_original_data %>%
  group_by(id, k) %>%
  summarise(ans = first(null)) -> eval_lineup_ans
```

```{r}
lineup_eval_dat %>%
  left_join(eval_lineup_ans, by = c("set" = "id", "plot" = "k")) %>%
  group_by(model, set) %>%
  summarise(n_sel = sum(!predict), detect = sum(ans == FALSE & predict == FALSE)) -> lineup_eval_summary
```

```{r}
cache_env <- new.env()
lineup_eval_summary %>%
  calc_p_value_multi(lineup_id = "set", detected = "detect", n_sel = "n_sel", cache_env = cache_env) -> cv_results
```
```{r}
cv_results %>%
  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
  left_join(select(eval_original_data, id, effect_size) %>%
              group_by(id) %>%
              summarise(effect_size = first(effect_size)),
            by = c("lineup_id" = "id")) -> cv_results

cv_results %>%
  mutate(type = ifelse(lineup_id <= 2500, "cubic", "heteroskedasticity")) -> cv_results
```

```{r}
merged_survey %>%
  mutate(exp_lineup_id = paste0(exp, "_", lineup_id)) %>%
  select(exp_lineup_id, type, exp, effect_size) %>%
  right_join(comb_pvalue) %>%
  mutate(n_eval = factor(n_eval)) %>%
  unnest_longer(col = p_value) %>%
  mutate(reject = eval_p_value(p_value, tol = 1e-2)) %>%
  mutate(reject = as.numeric(reject)) %>%
  ggplot() +
  geom_point(aes(log(effect_size), reject), alpha = 0.1, size = 1) +
  geom_smooth(aes(log(effect_size), reject, group = "conventional"),
              method = "glm", method.args = list(family = binomial), data = conv_result, 
              se = FALSE, size = 1, linetype = 2, col = "grey80") +
  geom_smooth(aes(log(effect_size), reject, group = n_eval, col = n_eval),
              method = "glm", method.args = list(family = binomial), size = 0.5, se = FALSE) +
  geom_smooth(aes(log(effect_size), as.numeric(reject), group = "cv"), data = cv_results,
              se = FALSE, size = 1, method = "glm", method.args = list(family = binomial)) +
  facet_wrap(~type, scales = "free_x") +
  labs(col = "Number of evaluations") +
  theme_light(base_size = 5) +
  theme(legend.position = "bottom") +
  scale_colour_viridis_d(option = "B", begin = 0.5, end = 0.9) +
  guides(col = guide_legend(nrow = 1, byrow=TRUE)) +
  xlab("Log of effect size") +
  ylab("Rejected")
```

### 2022-03-06 Thesis/_book/thesis.tex
add performance

\hypertarget{artificial-intelligence-four-approaches}{%
\section{Artificial intelligence: Four Approaches}\label{artificial-intelligence-four-approaches}}

@@

Considering there are different approches of AI, it is necessary to clarify the type of AI we aimed to build. Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research. We believed humanly behaviours was enough for demonstrating AI for the problem we concerned. Hence, adopting one of the capacities required by the Turing test, namely machine learning, along with

Thought process, human visual

\hypertarget{discussion-of-potential-methods}{%
\section{Discussion of potential methods}\label{discussion-of-potential-methods}}

@@

>>>>

\hypertarget{se:ai}{%
\section{Artificial intelligence: Four Approaches}\label{se:ai}}

@@

Considering there are different approches of AI discussed in \ref{se:ai}, it is necessary to clarify the type of AI we aimed to build. Understanding the human thought process or mimicking the human vision mechanism were not of interest in this research. We believed humanly behaviours was enough for demonstrating AI for the problem we concerned. Hence, two of the capacities required by the total Turing test, namely machine learning and computer vision will be adopted.

Another possible method was to define distance metrics to measure difference between data plots for making mathematically prefect decisions. This method would work if such metrics can be obtained manually or approximated algorithmically. However, we considered it was too complex as a starting point.

\hypertarget{automatic-visual-inference-system}{%
\section{Automatic visual inference system}\label{automatic-visual-inference-system}}

For an automatic visual inference system, there are several possibilities of the type of inputs.

Depend on the inputs and outputs, computer vision model can be integrated into an automatic visual inference system.

@@

\caption{\label{tab:null-true-eval}Performance of computer vision models on null residual plot and acutal data plot.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Model & Actual data plot correct rate & Null plot correct rate & Average correct rate\\
\midrule
VGG16 & 0.7387 & 0.8850 & 0.81185\\
VGG19 & 0.8029 & 0.8313 & 0.81710\\
Xception & 0.7370 & 0.9213 & 0.82915\\
ResNet152V2 & 0.6932 & 0.9233 & 0.80825\\
InceptionV3 & 0.6691 & 0.9399 & 0.80450\\
\bottomrule
\end{tabular}
\end{table}

\includegraphics{thesis_files/figure-latex/unnamed-chunk-44-1.pdf}
